Online Social Cohesion Reflects Real-World Group Action in Syria during the Arab Spring

### 2020-06-05 manuscript-chicago.tex
Submitted version

Correspondence concerning this article should be addressed to Megan Chiovaro, 406 Babbidge Road, Unit 1020, Storrs, CT 06269. E-mail: megan.chiovaro@uconn.edu}
\keywords{social cohesion, recurrence quantification analysis, complex systems, Twitter, Arab Spring\newline\indent Word count: X}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=english]{babel}
\else
  % load polyglossia as late as possible as it *could* call bidi if RTL lang (e.g. Hebrew or Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\fi

@@

\abstract{
In recent years, individuals involved in social movements have taken to social
media platforms to reach broader audiences at rapid speeds. Although
micro-blogging has become a primary means of information dissemination during
times of turmoil, the degree to which virtual mobilization reflects or drives
real-world movements is unclear. Here, we explore the dynamics of real-world
events and Twitter social cohesion in Syria during the Arab Spring. Using the
nonlinear methods cross-recurrence quantification analysis and windowed
cross-recurrence quantification analysis, we investigate if frequency of
events of different intensities are coupled with social cohesion found in
Syrian tweets. Results indicate that virtual social cohesion is coupled with
the counts of all and negative events each day but shows a more nuanced
connection to positive events, suggesting that virtual mobilization does
reflect real-world movements. We conclude with a discussion of implications
and applications of nonlinear methods in political science research.
}

@@

Prior to the advent of social media, people coordinated their protest behavior
using phone tree calling sheets and mimeographed flyers, through in-person
meetings of social organizations and churches, and through word of mouth. They
gathered together and hand-made protest signs that reflected their agendas. As
McAdam found, people were more likely to join movements when they had a personal
connection (McAdam, 1986; McAdam \& Paulsen, 1993): Historically, strong
ties have been strong predictors of involvement in high-risk protest activity.

The cohesiveness and strength of social movements in the pre-Internet era could
be operationalized by the number of people turning out to protest and through
the political events that precipitated and resulted from social gatherings \emph{en
masse}. Cohesive social movements were reflected in their cohesive placards and
posters. For example, the \enquote{I am a man} posters held by the men involved in the
1968 sanitation workers' strike reflected their internal cohesion and unified,
goal-oriented movement (Windsor, 2020).

However, the rise of the Internet and its decentralized online networks has
upended theories about recruitment to---and participation in---social movements.
Strong interpersonal ties have yielded to weak impersonal ties that strongly
predict involvement in social movements (Steinert-Threlkeld, 2017). Research
methodologies for measuring social movements have similarly evolved from
ethnographic interviews to \enquote{big data} analyses of language on social media.

Today, social media replicates and amplifies this phenomenon during moments of
social unrest. Amid the cacophony of messages during political upheaval---such
as the Arab Spring---consistent themes emerge through hashtags, retweets, and
the repeated mentions of concepts such as regime, security, demonstration, and
human rights. The cohesiveness of these themes (i.e., whether movement
participants are metaphorically singing in one voice) reflects the real-world
events happening contemporaneously with collective action in social movements or
civil conflicts.

@@

Disentangling the complicated relationship between mass online
mobilization---and social mobilization in person---continues to be relevant,
even as our modes of communication change. The political events that
transpired during and after the Arab Spring are still in the process of being
understood. As Zeitzoff (2017) recognizes, social media is
changing conflict, but the mechanisms and relationships by which it does so are
still largely unknown.

@@

The decades preceding the Arab Spring in Syria were marked by
social repression and fear. In 1971, Hafez al-Assad overthrew the previous
leader and remained in power for thirty years, presiding during
regional contentiousness and cooperation until his death in 2000. His son,
Bashar al-Assad continued the regime of authoritarian leadership as he assumed
his office. The Assad lineage hails from a minority Islamic sect (the Alawites)
in a multi-religious and multi-ethnic country comprising Christians, Sunni and
Shia Muslims, Kurds, and Palestinian refugees.

In December 2010, a Tunisian man named Mohamed Bouazizi was harassed and
unjustifiably stripped of his vendor cart by municipal officials. His
humiliation led him to set himself on fire in protest to the authoritarian rule,
ultimately leading to his death several days later. His actions served as a
catalyst for the Tunisian Revolution, and one week after his death, Zine El
Abidine Ben Ali---the longstanding authoritarian president---fled the country.
The events of Tunisia rapidly spread to surrounding countries also subject to
unjust authoritarian rule, including Yemen, Egypt, Libya, and Syria.

Following a wave of democratization efforts across the Middle East and North
Africa, modest Syrian protests began in March 2011. The regime responded harshly
with imprisonments, detentions, censorship, and military operations as the
resistance to Assad's regime grew and spread throughout the country. The
fighting escalated dramatically in March 2012, after increased unrest in the
western city of Homs. International actors---including the Red Cross/Red
Crescent Society, the United Nations General Assembly, the Security Council, and
individual world leaders---worked toward cease-fires, negotiations, and
concessions without much success. While many high-level military and diplomatic
officials defected from the regime, numerous loyal forces continued the regime's
repression, and the Syrian people turned to social media to help coordinate
their activism and communicate to the world. Between March and June 2012, more
than 8 million tweets were sent within Syria, revealing the details of the
conflict suppressed by the government.

Syria is a multi-ethnic, multi-lingual, and multi-religious state. Kurdish
people make up between 10-15\% of the population, Alawites between 8-15\%,
Christians about 10\%. There are also much smaller sects (e.g., Ismailis, Druze),
but Sunni Muslims comprise the vast majority of society. Most people speak
Arabic as their primary language. Most policy reports addressing the current
conflict report on the cohesion of the ruling Alawite military elite, rather
than on the cohesiveness of the opposition movement. However, the Sunni
majority's disaffection with the Syrian government and their lack of political
representation provide sufficient grievance and collective group identification
to make them a cohesive force.

@@

Massive online social cohesion can have real-world social manifestations. Social
media platforms provide one way to enable this kind of online social cohesion by
offering distributed information-gathering and real-time information
dissemination. Given its accessibility and proficiency as an information
exchange tool, Twitter has provided individuals involved in social movements an
opportunity to complement or even replace real-world social mobilization. For
example, during the 2012 Italian protests for global economic change, the riots
in Rome were mirrored by Twitter activity both on the day of the protests and in
the days following them (Vicari, 2013). Tweeting was determined to be more
effective locally in discussing real-world, real-time events than traditional
mainstream media platforms. As more and more individuals rely on social media to
obtain local news and information, it becomes increasingly important to
understand just how these micro-blogs reflect real-world action.

During the Arab Spring, social media provided real-time information about the
events transpiring in countries across the Middle East and North Africa. Our
understanding about the role of social media in social movements is still
evolving. Some argue that these mass movements would not have been possible
without efficient large-scale coordination, both physically on the ground and
virtually on social media platforms.

Along with Facebook, Twitter has been credited as essential to the Arab
Spring---both for protesters (as a means to coordinate with one another and to
communicate with the world) and for the government response (to monitor and
track protestors to crack down on dissent) (Comninos, 2011). The events of
Arab Spring have been thoroughly discussed within the political science
literature (e.g., Lutterbeck, 2013; Koehler, 2017), but to our knowledge, the direct \emph{coupling} between
Syrian Twitter cohesion and international action coordination has not been
investigated. In fact, some have been skeptical about the effect of social media
on collective action at all (Seay, 2014).

This skepticism can be summarized in a quote by Lisa Anderson:

\begin{quote}
The important story about the 2011
Arab revolts in Tunisia, Egypt, and Libya is not how the globalization of the
norms of civic engagement shaped the protesters' aspirations. Nor is it about
how activists used technology to share ideas and tactics. Instead, the critical
issue is how and why these ambitions and techniques resonated in their various
local contexts. (Anderson, 2011)
\end{quote}

\noindent Others have similarly expressed hesitation about the role of online
organizing. Some have argued that real-life strong ties outweigh online weak ties and
that the power of the strong ties through in-person contact is critical
to recruiting people for higher-risk activities
(Gladwell, 2010; McAdam, 1986).

Interestingly, these views may
actually help to make the case for bridging the gap between social media and
social mobilization. Such views suggest the strength of social ties in an online
environment is inextricably linked to the \emph{context} and \emph{ambitions} of the actors.
Moreover, research suggests that weak ties still retain considerable political
power: While strong ties tend to bind smaller groups, weak ties
tend to bind groups across distances (Granovetter, 1973).

Social media generates critical information about events, including whether and
how a demonstration will take place, how many people participate, and how those
in power (e.g., state, military, security apparatus) respond. Similarly, these
media platforms alert us to more global shifts within a conflict, such as when
fighting moves to a new city or region, the number of casualties and wounded,
and the types of weaponry used by parties to the conflict (Zeitzoff, 2017).
Social media is not entirely event-driven, however. In monitoring social media
channels, government forces may respond to the perception that citizens are
gaining strength and sympathy. Governments have been shown to be quite tolerant
of dissent so long as it does not metamorphose into collective action
(King, Pan, \& Roberts, 2013).

Citizens communicate with each other with increased transparency due to the
ubiquitous availability of social media such as low-cost and egalitarian
platforms like Twitter. Citizens are able to access these tools via mobile
phones and computers with Internet connectivity, providing a stream of real-time
shared information. Citizens can share the location of events like protests and
military incursions, the degree of state repression, and the number of
casualties. In turn, the government---interested in remaining in
power---monitors social media platforms to ascertain emerging threats
(Comninos, 2011), such as where new groups are forming, where existing
groups are strengthening, where gatherings are held, and who the idea
entrepreneurs are.

@@

\begin{quote}
I know that I have been away from the media for a long time, and I have missed
having direct contact with the citizens, but I have always been following up
with the daily occurrences and gathering the information so that my speech can
be built on what is said by the street.
\end{quote}

\noindent With few exceptions, Assad chose to keep social media platforms up and running.
Gohdes (2015) notes that regimes can attempt to thwart the growth
of social protest online and in the streets by shutting down the Internet.
During these times, however, she finds that military repression increases.
However, government information about the citizens decreases; as such, many
regimes have elected to maintain Internet connectivity as a means of monitoring
and surveilling their citizens.

At the same time, the insulation of online social platforms allows researchers
to detail real-world social dynamics at unprecedented resolution and scale
(Paxton \& Griffiths, 2017). The digital traces left by individuals and collectively as
groups are regarded as a valuable source of information that can offer powerful
insights into states of social disequilibrium, uprisings, and escalations into
civil wars. By examining these conversations, researchers can study the degree
of cohesion between speakers and the alignment between that cohesion and their
collective political behavior.

@@

People join social and political resistance movements because they feel a sense
of connection to the cause. Given that people are generally risk-averse, overcoming inertia
requires a strong connection to the group identity or a strong aversion to an
individual, event, or political injustice (Goldstone, 1980, 2011). Political resistance offers participants the
opportunity to \enquote{lose themselves} and relinquish their individual identity in
favor of identification with a larger group, contributing to a cohesive
group ready and mobilized for action (Hoffer, 2011).

Previous research on social movements supports the idea that participants are
more likely to join movements when they know and have a personal connection to
someone else already involved (McAdam, 1986). The type of collective
contentious behavior that manifests in society is a function of the available
political opportunity structures. The political opportunity structure are
\enquote{comprised of \emph{(sic)} specific configurations of resources, institutional
arrangements and historical precedents for social mobilization, which facilitate
the development of protest movements in some instances and constrain them in
others} (Kitschelt, 1986). What emerges as strategies and tactics in
contentious politics are, in essence, a function of what the system makes
possible. This includes the number of people and their disposition (
i.e.~nonviolent or violent), as well as their internal structure. The more cohesive a
group is, the greater threat they pose to the establishment.

Cohesion is also a familiar concept in the study of civil conflicts, as it
pertains to group affiliation and the ability to overcome collective action
problems. Individuals must act in concert---that is, cohesively---to
accomplish the task of challenging the authority of the state (Eisinger, 1973; Francisco, 1995; Olson, 1965; Tilly, 2004).
Recent work has considered rebel recruitment for participation in
civil conflicts as a function of internal cohesion (Gates, 2002):
Rebel movement success is influenced by the degree of group cohesion, which
itself is a function of geographic distance, ideology, and ethnicity.
Within Syria, multiple rebel groups exist, sourcing fighters locally
from the disaffected Syrian population and internationally through
fighters from transnational terrorist organizations. There is considerable
intergroup competition for participation in the rebel movement and
punishment for defectors for nonparticipation (Gates, 2002).
Foreign fighters would appear to be amply motivated to fight, given their
decision to relocate from a home country that may be located in the region
or much farther away.

Another factor influencing social cohesion during social movements and civil
wars is ethnic identity and the degree of ethnic conflict associated with the
movement. Greater degrees of ethnic conflict increase the salience of ethnic
identity, leading people to identify more strongly with their own ethnic
identity (Sambanis \& Schulhofer-Wohl, 2009). As ethnic conflict intensifies, groups seek to become
more quintessentially like the typical member of their ethnic group. This can
manifest in a number of ways, including in behavior on social media. However,
this may also include a general identification across ethnic groups in a
membership of an opposition to an oppressive leader.

The civil war literature has further explored the role of social cohesion in
detail. Social homogeneity increases the likelihood of social conflict,
especially when the society is polarized (Collier \& Hoeffler, 2004); this is partly
due to nationalist groups' strong identification with a collective identity
(Cederman \& Girardin, 2007) and their social solidarity (Sambanis \& Schulhofer-Wohl, 2009). These
same processes can be similarly important for rebels (Anyanwu, 2002).
Referencing Gurr's (2000) work on minorities at
risk, Regan and Norton (2005) note that

\begin{quote}
a religious or linguistic minority
might suffer disproportionately in a given society, and this form of
aggrievement can lead to unrest across the social lines that distinguish the
minority group. Therefore, while collective cohesion is a function of targeted
maldistributions, it is individual penury that allows for this collective
mobilization. (Regan \& Norton, 2005)
\end{quote}

@@

Just as real-world social cohesion can be quantified, so can online cohesion.
One kind of cohesion is the similarity in language use on social
media. This implicit or explicit similarity can indicate shared
beliefs, thoughts, and feelings, as has been studied in research on language
style in psycholinguistics and social psychology (Brennan \& Clark, 1996; Gonzales, Hancock, \& Pennebaker, 2010; Tausczik \& Pennebaker, 2010). Although previous work on language style has often
focused on individuals and small groups, this can be scaled up to examine
massive social-level changes (Fusaroli et al., 2015; Michel et al., 2011).

The words that comprise social media posts are the first level of cohesion that
we will consider. In Figure \ref{fig:cohesion-in-tweets}, we see four
sequential tweets, each containing four words. Each sequential tweet contains
some content included in the previous tweets. The linguistic construct of
content word overlap measures the relationship across units of meaning---here,
tweets---where identical and semantically similar words are repeated
(Graesser, McNamara, Louwerse, \& Cai, 2004). These repeated lexical items constitute a theme; on Twitter,
this might be a particular hashtag, a name, a location, or other nominal
category. This localized cohesion is referential; words are repeated and
co-referred across tweets, creating a sense of continuity and cohesiveness of
meaning and message.

Typically, these analyses focus on \emph{content words} (e.g., nouns, adjectives,
verbs), which convey more meaningful semantic information than \emph{function words}
(e.g., articles, conjunctions). While function words can provide important
information about complexity and language structure
(Tausczik \& Pennebaker, 2010), content words have more specialized meaning and
are therefore more useful when researchers want to understand \emph{what} people are
discussing, rather than \emph{how} they are discussing it (Graesser et al., 2004; McNamara, Graesser, McCarthy, \& Cai, 2014).

In estimating cohesion in Twitter posts, tweets that share content words
will yield higher cohesion values, even among speakers with opposing
views on the subject matter. Repetition of terms and concepts is captured by
reiteration, which is a type of lexical cohesion in which a subsequent item
repeats a previous item, or refers to it through synonymous language
(Halliday \& Hasan, 1976). In social media, we can conceive of retweets (i.e., reposts
or forwards of previous messages) as a form of reiteration
that increase message cohesion. Since retweets replicate verbatim some part of
another tweet, these tweets would also have higher cohesion. Hashtags function
similarly and are included in our analysis.

\begin{figure}
\includegraphics[width=1\linewidth]{./figs/cohesion-in-tweets} \caption{A visual representation of cohesion in tweets. Content word overlap indicates a shared relationship among tweets. The shared words--either identical or semantically similar--constitute a theme around which cohesion is occuring.}\label{fig:cohesion-in-tweets}
\end{figure}

The use of similar language can provide insight into \emph{interlocutor cohesion}, or
the degree to which people are talking about similar things. This can---as noted
earlier---include people who are talking about similar things from radically
perspectives, as well as those who are talking about something from the same
perspective. Previous research from political science has investigated these
ideas (Windsor, 2020): The \enquote{Occupy Wall Street} movement has been held up as an
example of a movement with low interlocutor and linguistic cohesion because of
its wide range of signs and slogans, while the \enquote{I am a man} protest (captured by
Ernest Withers during the 1968 sanitation workers' strike in Memphis, TN) or the
\enquote{Black Lives Matter} placards (following the uprising in Ferguson, MO) have high
lexical and interlocutor cohesion (Edgar \& Johnson, 2018).

Fed by lexical cohesion, interlocutor cohesion ebbs and flows along a timeline
contextualized by variation in real world events. Interlocutor cohesion may
increase preceding an in-person social mobilization event, alerting a repressive
government that collective mobilization is happening. Interlocutor cohesion may
also respond to exogenous events such as spontaneous, unprovoked militarized
conflict. If government forces target a new location using military weapons,
this can precipitate online mobilization and discussion of the event.

Given the importance of lexical and interlocutor cohesion for messaging in
political movements, lexical cohesion may be an important bit of trace data for
researchers interested in leveraging social media data to analyze social
movements. Similar to large-scale studies of sociocultural movements (sometimes
referred to as \emph{culturomics}; e.g., Michel et al., 2011), political
scientists can track the ways in which conversations on social media coalesce or
diffuse around similar content words. In this way, it matters relatively little
whether the interlocutors have precisely the same views: The point is to
understand whether the society-level conversation is converging on similar
language.

@@

We tie the constructs of lexical, interlocutor, and social cohesion together in
a theoretical framework called \emph{sociolinguistic cohesion}, as shown in Figure
\ref{fig:theoretical-framework}. Several interconnected mechanisms are in
motion here. Previous values of linguistic cohesion should influence future
values of linguistic cohesion and future actions in the real world space.
Previous events influence future events and future linguistic cohesion in
social media.

>>>>

Correspondence concerning this article should be addressed to Megan Chiovaro, 406 Babbidge Road, Unit 1020, Storrs, CT 06269. E-mail: megan.chiovaro@uconn.edu}
\keywords{social cohesion, recurrence quantification analysis, complex systems, Twitter, Arab Spring\newline\indent Word count: X}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{csquotes}
\usepackage{setspace}
\ifxetex
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[shorthands=off,main=english]{babel}
\fi

@@

\abstract{
In recent years, political activists have taken to social media platforms to
rapidly reach broad audiences. Despite the prevalence of micro-blogging in
these sociopolitical movements, the degree to which virtual mobilization
reflects or drives real-world movements is unclear. Here, we explore the
dynamics of real-world events and Twitter social cohesion in Syria during the
Arab Spring. Using the nonlinear methods cross-recurrence quantification
analysis and windowed cross-recurrence quantification analysis, we investigate
if frequency of events of different intensities are coupled with social
cohesion found in Syrian tweets. Results indicate that virtual social cohesion
is coupled with the counts of all and negative events each day but shows a
more nuanced connection to positive events, suggesting that virtual
mobilization does reflect real-world movements. We conclude with a discussion
of implications and applications of nonlinear methods in political science
research.
}

@@

Before social media, people coordinated protests using phone tree calling sheets
and mimeographed flyers, through in-person meetings of social organizations and
churches, and by word of mouth. They gathered together and hand-made protest
signs that reflected their agendas. As McAdam found, people were more likely to
join movements when they had a personal connection (McAdam 1986; McAdam and Paulsen 1993): Historically, strong ties have been strong predictors of
involvement in high-risk protest activity.

The cohesiveness and strength of social movements before the Internet were
operationalized by the number of people present and through the political events
that precipitated and resulted from social gatherings \emph{en masse}. Cohesive
social movements were reflected in their cohesive placards and posters. For
example, the pervasive \enquote{I am a man} posters during the 1968 sanitation workers'
strike reflected their internal cohesion and unified, goal-oriented movement
(Windsor 2020).

However, the rise of the Internet and its decentralized online networks have
upended theories about recruitment to---and participation in---social movements.
Strong interpersonal ties have yielded to weak impersonal ties that strongly
predict involvement in social movements (Steinert-Threlkeld 2017). Measurements
of social movements have similarly evolved from ethnographic interviews to \enquote{big
data} analyses of social media.

Amid the cacophony of messages during political upheaval---such
as the Arab Spring---consistent themes emerge through hashtags, retweets, and
the repeated mentions of concepts such as regime, demonstration, and
human rights. The cohesiveness of these themes (i.e., whether movement
participants are metaphorically singing in one voice) reflects the real-world
events happening contemporaneously with collective action in social movements or
civil conflicts.

@@

Disentangling the complicated relationship between mass online
mobilization---and social mobilization in person---continues to be relevant,
even as our modes of communication change. Accordingly, the political events
around the Arab Spring are still being understood. As Zeitzoff
(2017) recognizes, social media is changing conflict, but the
mechanisms and relationships by which it does so are still largely unknown.

@@

The decades preceding the Arab Spring in Syria were marked by repression and
fear. In 1971, Hafez al-Assad overthrew the previous leader and remained in
power for thirty years, presiding during regional contentiousness and
cooperation until his death in 2000. His son, Bashar al-Assad, continued the
regime of authoritarian leadership after assuming his office. The Assad lineage
hails from a minority Islamic sect (the Alawites) in a multi-religious and
multi-ethnic country comprising Christians, Sunni and Shia Muslims, Kurds, and
Palestinian refugees.

In December 2010, a Tunisian man named Mohamed Bouazizi was unjustifiably
stripped of his vendor cart by municipal officials. His humiliation led him to
set himself on fire in protest to the authoritarian rule, leading to his death
several days later. His actions served as a catalyst for the Tunisian
Revolution, and one week after his death, Zine El Abidine Ben Ali---the
longstanding authoritarian president---fled the country. The events of Tunisia
rapidly spread to surrounding countries also subject to unjust authoritarian
rule, including Egypt, Libya, and Syria.

Following a wave of democratization efforts across the Middle East and North
Africa, modest Syrian protests began in March 2011. The regime responded harshly
with imprisonments, detentions, censorship, and military operations as the
resistance to Assad's regime spread. The fighting escalated dramatically in
March 2012, after increased unrest in the western city of Homs. International
actors---including the Red Cross/Red Crescent Society, the United Nations
General Assembly, the U.N. Security Council, and individual world
leaders---worked toward cease-fires, negotiations, and concessions without much
success. While many high-level military and diplomatic officials defected,
numerous loyal forces continued the regime's repression, and the Syrian people
turned to social media to coordinate their activism and communicate to the
world. Between March and June 2012, more than 8 million tweets were sent within
Syria, revealing the details of the conflict suppressed by the government.

Syria is a multi-ethnic, multi-lingual, and multi-religious state. Kurdish
people make up between 10-15\% of the population, Alawites between 8-15\%, and
Christians about 10\%. There are also much smaller sects (e.g., Ismailis, Druze),
but Sunni Muslims comprise the vast majority of society. Most people speak
Arabic as their primary language. Most policy reports addressing the current
conflict report on the cohesion of the ruling Alawite military elite, rather
than on the cohesiveness of the opposition movement. However, the Sunni
majority's disaffection with the Syrian government and their lack of political
representation provide sufficient grievance and collective group identification
to make them a cohesive force.

@@

Massive online social cohesion can manifest in real-world social behavior.
Social media platforms enable online social cohesion by offering distributed
information-gathering and real-time information dissemination. Given its
accessibility and proficiency as an information exchange tool, Twitter has
provided social and political activists an opportunity to complement or even
replace real-world social mobilization. During the 2012 Italian protests for
global economic change, tweeting was determined to be more effective locally in
discussing real-world, real-time events than traditional mainstream media
platforms (Vicari 2013). As people increasingly rely on social media for
local news and information, we must understand exactly how these micro-blogs
connect to and reflect real-world action.

Along with Facebook, Twitter has been credited as essential to the Arab
Spring---both for protesters and for the government's response to them
(Comninos 2011). The Arab Spring has been deeply examined within the
political science literature (e.g., Lutterbeck 2013; Koehler 2017), but to our knowledge, the direct \emph{coupling} between
Syrian Twitter cohesion and international action has not been investigated. In
fact, some have been skeptical about the effect of social media on collective
action at all (Seay 2014).

This skepticism is summarized in a quote by Lisa Anderson (2011):

\begin{singlespace}
\setlength{\leftskip}{1cm}

\noindent The important story about the 2011
Arab revolts in Tunisia, Egypt, and Libya is not how the globalization of the
norms of civic engagement shaped the protesters' aspirations. Nor is it about
how activists used technology to share ideas and tactics. Instead, the critical
issue is how and why these ambitions and techniques resonated in their various
local contexts.

\setlength{\leftskip}{0pt}
\end{singlespace}

\noindent Others have similarly expressed hesitation about the role of online
organizing. Some have argued that real-life strong ties outweigh online weak
ties and that the power of the strong ties through in-person contact is critical
to recruiting people for higher-risk activities (Gladwell 2010; McAdam 1986).

Interestingly, these views may help make the case for bridging social media and
social mobilization. Such views suggest the strength of online social ties is
inextricably linked to the \emph{context} and \emph{ambitions} of the actors. Moreover,
research suggests that weak ties still retain considerable political power:
While strong ties tend to bind smaller groups, weak ties tend to bind groups
across distances (Granovetter 1973).

Social media generates critical information about events, including whether and
how a demonstration will take place, how many people participate, and how those
in power (e.g., state, military) respond. Similarly, these media platforms alert
us to more global shifts within a conflict, such as when fighting moves to a new
city or region, the number of casualties and wounded, and the types of weaponry
used by parties (Zeitzoff 2017). In monitoring social media channels,
government forces may respond to the perception that citizens are gaining
strength and sympathy. Governments have been shown to be quite tolerant of
dissent so long as it does not metamorphose into collective action
(King, Pan, and Roberts 2013).

Citizens communicate with each other with increased transparency due to the
ubiquitous availability of social media such as low-cost and egalitarian
platforms like Twitter. Citizens are able to access these tools via mobile
phones and computers with Internet connectivity, providing a stream of real-time
shared information. Citizens can share the location of events like protests and
military incursions, the degree of state repression, and the number of
casualties. In turn, the government---interested in remaining in
power---monitors social media platforms to ascertain emerging threats
(Comninos 2011), such as where new groups are forming, where existing
groups are strengthening, where gatherings are held, and who the leaders are.

@@

\begin{singlespace}
\setlength{\leftskip}{1cm}

\noindent I know that I have been away from the media for a long time, and I
have missed having direct contact with the citizens, but I have always been
following up with the daily occurrences and gathering the information so that my
speech can be built on what is said by the street.

\setlength{\leftskip}{0pt}
\end{singlespace}

\noindent With few exceptions, Assad chose to keep social media platforms up and
running. Gohdes (2015) notes that regimes can attempt to thwart
the growth of social protest online and in the streets by shutting down the
Internet. During these times, military repression increases, but government
information about the citizens decreases. As a result, regimes may elect to
maintain Internet connectivity to monitor and surveil their citizens.

At the same time, the insulation of online social platforms allows researchers
to detail real-world social dynamics at unprecedented resolution and scale
(Paxton and Griffiths 2017). The digital traces left by individuals and collectively as
groups are regarded as a valuable source of information that can offer powerful
insights into states of social disequilibrium, uprisings, and escalations into
civil wars. By examining these conversations, researchers can study the degree
of cohesion between speakers and the alignment between that cohesion and their
collective political behavior.

@@

People join social and political resistance movements because they feel a sense
of connection to the cause. Given that people are generally risk-averse,
overcoming inertia requires a strong connection to the group identity or a
strong aversion to an individual, event, or political injustice
(Goldstone 1980, 2011). Political resistance offers participants the
opportunity to \enquote{lose themselves} and relinquish their individual identity in
favor of identification with a larger group, contributing to a cohesive group
ready and mobilized for action (Hoffer 2011).

The type of collective contentious behavior that manifests in society is a
function of the available political opportunity structures. The political
opportunity structure is \enquote{comprised of \emph{(sic)} specific configurations of
resources, institutional arrangements and historical precedents for social
mobilization, which facilitate the development of protest movements in some
instances and constrain them in others} (Kitschelt 1986). What emerges
as strategies and tactics in contentious politics are a function of
what the system makes possible. This includes the number of people and their
disposition (i.e.~nonviolent or violent), as well as their internal structure.
The more cohesive a group is, the greater threat they pose to the establishment.

Cohesion is a familiar concept in the study of civil conflicts, as it pertains
to group affiliation and the ability to overcome collective action problems.
Individuals must act in concert---that is, cohesively---to accomplish the task
of challenging state authority (Olson 1965; Eisinger 1973; Tilly 2004; Francisco 1995).
Viewing rebel recruitment during civil conflicts as a function of internal
cohesion, rebel movement success is influenced by the degree of group cohesion,
which itself is a function of geographic distance, ideology, and ethnicity
(Gates 2002). Within Syria, multiple rebel groups exist, sourcing
fighters locally (from the disaffected Syrian population) and internationally
(from transnational terrorist organizations). Groups compete intensely with one
another for recruitment in the rebel movement and often punish defectors for
nonparticipation (Gates 2002).

Another factor influencing social cohesion during social movements and civil
wars is ethnic identity and the degree of ethnic conflict associated with the
movement. Greater degrees of ethnic conflict increase the salience of ethnic
identity, leading people to identify more strongly with their own ethnic
identity (Sambanis and Schulhofer-Wohl 2009). As ethnic conflict intensifies, groups seek to become
more quintessentially like the typical member of their ethnic group. This can
manifest in a number of ways, including in behavior on social media. However,
this may also include a general identification across ethnic groups in a
membership of an opposition to an oppressive leader.

The civil war literature has also explored the role of social cohesion. Social
homogeneity increases the likelihood of social conflict, especially when the
society is polarized (Collier and Hoeffler 2004); this is partly due to nationalist
groups' strong identification with a collective identity (Cederman and Girardin 2007)
and their social solidarity (Sambanis and Schulhofer-Wohl 2009). These same processes can be
similarly important for rebels (Anyanwu 2002). Referencing Gurr's
(2000) work on minorities at risk, Regan and Norton
(2005) note that

\begin{singlespace}
\setlength{\leftskip}{1cm}

\noindent a religious or linguistic minority
might suffer disproportionately in a given society, and this form of
aggrievement can lead to unrest across the social lines that distinguish the
minority group. Therefore, while collective cohesion is a function of targeted
maldistributions, it is individual penury that allows for this collective
mobilization.

\setlength{\leftskip}{0pt}
\end{singlespace}

@@

Just as real-world social cohesion can be quantified, so can online cohesion.
One kind of cohesion is the similarity in language use on social
media. This implicit or explicit similarity can indicate shared
beliefs, thoughts, and feelings, as has been studied in research on language
style in psycholinguistics and social psychology (Gonzales, Hancock, and Pennebaker 2010; Tausczik and Pennebaker 2010; Brennan and Clark 1996). Although previous work on language style has often
focused on individuals and small groups, this can be scaled up to examine
massive social-level changes (Michel et al. 2011; Fusaroli et al. 2015).

The first level of cohesion that we will consider are the words within social
media posts. In Figure \ref{fig:cohesion-in-tweets}, we see four sequential
tweets of four words each. Each sequential tweet contains some content
included in the previous tweets. The linguistic construct of content word
overlap measures the relationship across units of meaning---here, tweets---where
identical and semantically similar words are repeated (Graesser et al. 2004). These
repeated lexical items constitute a theme; on Twitter, this might be a
particular hashtag, a name, a location, or another nominal category. This
localized cohesion is referential; words are repeated and co-referred across
tweets, creating a sense of continuity and cohesiveness.

Typically, these analyses focus on \emph{content words} (e.g., nouns, adjectives,
verbs), which convey more meaningful semantic information than \emph{function words}
(e.g., articles, conjunctions). While function words can provide important
information about complexity and language structure
(Tausczik and Pennebaker 2010), content words have more specialized meaning and
are therefore more useful when researchers want to understand \emph{what} people are
discussing, rather than \emph{how} they are discussing it (Graesser et al. 2004; McNamara et al. 2014).

In estimating cohesion in Twitter posts, tweets that share content words
yield higher cohesion values, even among speakers with opposing
views on the subject matter. Repetition of terms and concepts is captured by
reiteration, which is a type of lexical cohesion in which a subsequent item
repeats a previous item, or refers to it through synonymous language
(Halliday and Hasan 1976). In social media, we can conceive of retweets (i.e., reposts
or forwards of previous messages) as a form of reiteration
that increase message cohesion. Since retweets replicate verbatim some part of
another tweet, these tweets would also have higher cohesion. Hashtags function
similarly and are included in our analysis.

\begin{figure}
\includegraphics[width=1\linewidth]{./figs/cohesion-in-tweets} \caption{A visual representation of cohesion in tweets. The shared content words--either identical or semantically similar--constitute a theme around which cohesion is occuring.}\label{fig:cohesion-in-tweets}
\end{figure}

The use of similar language can provide insight into \emph{interlocutor cohesion}, or
the degree to which people are talking about similar things. This can include
people who are talking about similar things from radically perspectives, as well
as those who are talking about something from the same perspective. Previous
research from political science has investigated these ideas (Windsor 2020): The
\enquote{Occupy Wall Street} movement has been held up as an example of a movement with
low interlocutor and linguistic cohesion because of its wide range of signs and
slogans, while the \enquote{I am a man} protest (during the 1968 sanitation workers'
strike in Memphis, TN) or the \enquote{Black Lives Matter} placards (following the
uprising in Ferguson, MO) have high lexical and interlocutor cohesion
(Edgar and Johnson 2018).

Fed by lexical cohesion, interlocutor cohesion ebbs and flows, contextualized by
changing real world events. Interlocutor cohesion may increase preceding an
in-person social mobilization event, alerting a repressive government that
collective mobilization is happening. Interlocutor cohesion may also respond to
exogenous events such as spontaneous, unprovoked militarized conflict.

Given the importance of lexical and interlocutor cohesion for messaging in
political movements, lexical cohesion may be important trace data for
researchers interested in leveraging social media data to analyze social
movements. Similar to large-scale studies of sociocultural movements (sometimes
referred to as \emph{culturomics}; e.g., Michel et al. 2011), political
scientists can track the ways in which conversations on social media coalesce or
diffuse around similar content words. In this way, it matters relatively little
whether the interlocutors have precisely the same views: The point is to
understand whether the society-level conversation is converging on similar
themes.

@@

We tie the constructs of lexical, interlocutor, and social cohesion together in
a theoretical framework called \emph{sociolinguistic cohesion}, as shown in Figure
\ref{fig:theoretical-framework}. Previous values of linguistic cohesion should influence future values of linguistic cohesion and future actions in the real world space.
Previous events influence future events and future linguistic cohesion in
social media.
